---
title: "Analise Conjunta Virginia"
author: "Chris Simoes"
date: "3/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_chunk$set(message=FALSE)
knitr::opts_chunk$set(cache=FALSE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(results="hold")
knitr::opts_chunk$set(collapse=TRUE)
knitr::opts_chunk$set(fig.show="hold")
knitr::opts_chunk$set(tidy.opts = list(wrap=FALSE))
```

## Objectives
1. Find the best candidate to become a commercial variety
2. Develop a selection index

### Script setup
Install packages, start the README.md, set the random seed  
```{r Script setup, echo=FALSE, include=FALSE}
ip <- installed.packages()
packages_used <- c("agricolae","emmeans","lme4","tidyverse", "devtools", "here")

ip <- installed.packages()
for (package in packages_used){
  if (!(package %in% ip[,"Package"])) install.packages(package)
}#END packages_used

here::i_am("analysis/analiseConjuntaVA.Rmd")

source(here::here("code/addToREADME.R"))

addToREADME(paste0("## ", rmarkdown::metadata$title), append=F)
addToREADME(c(date(), ""))

packages_info <- ip[packages_used, c("Package", "Version", "Built")]
addToREADME(c("The packages used in this script are:", "Package, Version, Built"))
apply(packages_info, 1, function(vec) addToREADME(paste(vec, collapse=" ")))
addToREADME("")

random_seed <- 835246
set.seed(random_seed)
addToREADME(paste("The random seed is", random_seed))
```

```{r import data, echo=TRUE}
va <- read.csv("data/A427_converted.csv", sep=",", header=T)
name_cols <- c("germplasmName","Replication","Location","kgHa","DollarKg","DollarHa","IQS","Talo","TotalSuggar","TotalAlcaloids")
colnames(va) <- name_cols

Traits <- colnames(va)[4:length(va)]
nTraits <- ncol(va)-3
Treatments <- unique(va$germplasmName)
Locations <- unique(va$Location)

#This is the economic value for each trait
econValue <- c(1.5,1.01,1.45,1.55,-1.25,-1.1,-1.16)


pre_result <- matrix(data=NA, nrow = nTraits, ncol = 3)
colnames(pre_result)<- c("h2", "varG", "varP")
rownames(pre_result)<-Traits

library("lme4")
library("emmeans")

#Centering the data - Function scale sets mean = 0 
#It is important to scale the data in order to get a sum of different traits afterwards
va_centered <- va

for(i in 1:nTraits){
  va_centered[,3+i] <- scale(va[,i+3], scale = TRUE)
}

for (i in 4:length(va_centered)){
  
  Pheno <- colnames(va_centered)[i]
  model <- lmer( get(Pheno) ~ (1|germplasmName) + (1|Replication) +(1|Location) + (1|germplasmName:Location),
                 na.action = na.exclude,
                 data=va_centered)
  
  variance = as.data.frame(VarCorr(model))
  gvar <- variance[2,"vcov"]
  envar <- variance[3, "vcov"]
  resvar <- variance [5, "vcov"]
  addCor <- variance[2,"sdcor"]
  phenovar <- gvar + envar + resvar
  H2 = gvar/phenovar
  pre_result[i-3,1] <- H2
  pre_result[i-3,2] <- gvar
  pre_result[i-3,3] <- phenovar
  
}

preresults_table <- knitr::kable(pre_result)
preresults_table

```

## Pre processing

Preparing for Pheno and Geno matrix
Calculating regression coefficients 

```{r pre process, echo = TRUE}

##regression coefficient - if necessary
# rP <- c()
# count=1
# for(i in 1:nTraits){
#   t1 <- as.numeric(va[,i+3])
#   var1 <- pre_result[i,3]
#   if(i < nTraits){
#     for (j in (i+1):nTraits){
#       t2 <- as.numeric(va[,j+3])
#       var2 <- pre_result[j,3]
#       cal_rP <- var(t1,t2, na.rm = TRUE)/sqrt(var1*var2)
#       rP[count] <- cal_rP
#       count=count+1
#     }
#   }
# }

# 1 Phenotypic matrix
#The of diagonal is defined by the covariation between the traits.
# cov(1,2) = rP x sqrt(varPheno[1] x varPheno[2])
PhenoMatrix <- matrix(data = NA, ncol = nTraits, nrow = nTraits)
for (i in 1:nTraits){
  t1 <- as.numeric(va_centered[,i+3])
  if (i < nTraits){
    for (j in (i+1):nTraits){
      t2 <- as.numeric(va_centered[,j+3])
      offDiagonal <- var(t1,t2, na.rm = TRUE)
      PhenoMatrix[i,j] <- offDiagonal
      PhenoMatrix[j,i] <- offDiagonal
    }
  }
  PhenoMatrix[i,i] <- pre_result[i,3]
}


# 2 Genotypic matrix
GenoMatrix <- diag(c(pre_result[,2]))


# 3 Economic value
# Already defined as econValue

# Calculate the Index. It is a matrix with 1 column and nTraits rows
SelectionIndex <- solve(PhenoMatrix) %*% GenoMatrix %*% econValue
rownames(SelectionIndex) <- Traits
colnames(SelectionIndex) <- "Index"

print(SelectionIndex)

```


## Making Selection with Index

Here I'm using the index to multiply by each trait. After that it will be possible make a rank.

```{r making selection, echo=TRUE}
library(dplyr)
library(tidyr)

#Calculate the index per germplasmName
preresults <- va_centered %>% group_by(germplasmName) %>% summarise(kgHa = mean(kgHa, na.rm=TRUE),
                                                                   DollarKg = mean(DollarKg, na.rm=TRUE),
                                                                   DollarHa = mean(DollarHa, na.rm=TRUE),
                                                                   IQS = mean(IQS, na.rm=TRUE),
                                                                   Talo = mean(Talo, na.rm=TRUE),
                                                                   TotalSuggar = mean(TotalSuggar, na.rm=TRUE),
                                                                   TotalAlcaloids = mean(TotalAlcaloids, na.rm=TRUE)
                                                                   )

#Removing extra column with germplasmName
preresults <- data.matrix(subset(preresults, select= -c(germplasmName)))

#Naming rows and cols
colnames(preresults) <- Traits
rownames(preresults) <- Treatments

#Index Matrix has 1 column and the number of rows is the number of germplasmName
index_matrix <- preresults %*% SelectionIndex

# Creating a matrix with real average (not centered)
preresults2 <- va %>% group_by(germplasmName) %>% summarise(kgHa = mean(kgHa, na.rm=TRUE),
                                                                   DollarKg = mean(DollarKg, na.rm=TRUE),
                                                                   DollarHa = mean(DollarHa, na.rm=TRUE),
                                                                   IQS = mean(IQS, na.rm=TRUE),
                                                                   Talo = mean(Talo, na.rm=TRUE),
                                                                   TotalSuggar = mean(TotalSuggar, na.rm=TRUE),
                                                                   TotalAlcaloids = mean(TotalAlcaloids, na.rm=TRUE)
                                                                   )

preresults2 <- data.matrix(subset(preresults2, select= -c(germplasmName)))
 
#print(index_matrix)
# Result matrix has all traits + selection index
results<-as.data.frame(cbind(preresults2, index_matrix))
results <-results[order(results$Index, decreasing = TRUE),]

#Print the result with Rmarkdown 
results_table <- knitr::kable(results)
results_table

```

## Stability analysis with AMMI
This is using Bilinear sowftware

```{r stability, echo=TRUE}

library("Bilinear")

# # Prepare a matrix with mean of treatments per location
preStability <- data.matrix(va %>% group_by(Location, germplasmName) %>% summarise(kgHa = mean(kgHa, na.rm=TRUE),
                                                                   DollarKg = mean(DollarKg, na.rm=TRUE),
                                                                   DollarHa = mean(DollarHa, na.rm=TRUE),
                                                                   IQS = mean(IQS, na.rm=TRUE),
                                                                   Talo = mean(Talo, na.rm=TRUE),
                                                                   TotalSuggar = mean(TotalSuggar, na.rm=TRUE),
                                                                   TotalAlcaloids = mean(TotalAlcaloids, na.rm=TRUE)
                                                                   ))

Stability <- preStability[,3:9] %*% SelectionIndex
Stability <- cbind(preStability[,2], preStability[,1], Stability)
colnames(Stability) <- c("germplasmName","Location", "Index")
rownames(Stability) <- preStability[,2]

Stability <- tapply(Stability[,3], list (Stability[,1], Stability[,2]), mean)
AMMI_Index <- bilinear(x = Stability , verbose=F)

#Stability using a specific trait
Trt <- tapply(va$DollarKg, list (va$germplasmName, va$Location), mean)
AMMI_Trt <- bilinear(x = Trt , verbose=F)
AMMIplot(AMMI_Trt)
# abline(v=-0.05, col="red")
# abline(v=+0.05, col="red")
# AMMIplot(AMMIfit, PC=2)

library("agricolae")
## Exemple: mmodel<- with(va,AMMI(Location, germplasmName, Replication, DollarHa, console=FALSE))
Stability_AMMI <- data.matrix(va_centered[,4:10]) %*% SelectionIndex
va_centered <- cbind(va_centered, Stability_AMMI)
colnames(va_centered) <- c(name_cols,"Index")

model<- with(va_centered,AMMI(Location, germplasmName, Replication, Index, console=TRUE))
model$ANOVA
# # see help(plot.AMMI)
# # biplot
plot(model)
AMMI.contour(model,distance=0.398,shape=8,col="red",lwd=2,lty=5)
# # triplot PC 1,2,3 
# plot(model, type=2, number=TRUE)
# # biplot PC1 vs Yield 
# plot(model, first=0,second=1, number=TRUE)
# # Example 2


```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
